---
title: "Homework 9"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
library(rethinking)
library(ggplot2)
library(dplyr)
library(tidyr)
library(dagitty)
library(ggdag)
library(rstan)

```

## Question 1

Revisit the Bangladesh fertility data, data(bangladesh). Fit a model with both
varying intercepts by district_id and varying slopes of urban (as a 0/1 indicator
variable) by district_id. You are still predicting use.contraception. Inspect
the correlation between the intercepts and slopes. Can you interpret this correlation,
in terms of what it tells you about the pattern of contraceptive use in the sample? It
might help to plot the varying effect estimates for both the intercepts and slopes,
by district. Then you can visualize the correlation and maybe more easily think
through what it means to have a particular correlation. Plotting predicted proportion of women using contraception, in each district, with urban women on one axis
and rural on the other, might also help.

```{r}

data(bangladesh)

d_q1 <- bangladesh

d_q1$district_id <- as.integer(as.factor(d_q1$district))

dat_q1 <- data.frame(
  use.contraception=d_q1$use.contraception,
  district_id = as.factor(d_q1$district),
  urban=d_q1$urban
)

dat_list_q1 <- list(
  C=d_q1$use.contraception,
  D=d_q1$district_id,
  U=d_q1$urban,
  Uid=d_q1$urban + 1L,
  N=1934,
  K=max(d_q1$district_id)
)

```


Lets first fit the fixed effect model using MLE GLM model

```{r}


q1_a <- glm(use.contraception ~ district_id : urban, family=binomial, data=dat_q1)

summary.glm(q1_a, correlation = FALSE)

```

Lets fit the fixed effect model using bayesian inference

```{r}

model_q1_1_fit <- stan(file='week09/09_q1_1.stan', data=dat_list_q1, cores=4)
loo(model_q1_1_fit)
```


```{r}
model_q1_2_fit <- stan(file='week09/09_q1_2.stan', data=dat_list_q1, cores=4)
loo(model_q1_2_fit)
```

```{r}
model_q1_3_fit <- stan(file='week09/09_q1_3.stan', data=dat_list_q1, cores=4)
loo(model_q1_3_fit)
```

```{r}
model_q1_4_fit <- stan(file='week09/09_q1_4.stan', data=dat_list_q1, cores=4)
loo(model_q1_4_fit)
```

## Question 2

Now consider the predictor variables age.centered and living.children,
also contained in data(bangladesh). Suppose that age influences contraceptive
use (changing attitudes) and number of children (older people have had more time
to have kids). Number of children may also directly influence contraceptive use.
Draw a DAG that reflects these hypothetical relationships. Then build models needed
to evaluate the DAG. You will need at least two models. Retain district and urban, as in Problem 1. What do you conclude about the causal influence of age and
children?

Lets first decide on the DAG.


```{r}
dag <- dagitty("dag{
A -> N
A -> C
N -> C
}")

adjustmentSets( dag , exposure="A" , outcome="C" , effect="direct" )

```
```{r}
plot(dag)
```


```{r}
d_q2 <- bangladesh

d_q2$district_id <- as.integer(as.factor(d_q2$district))

dat_q2 <- data.frame(
  use.contraception=d_q2$use.contraception,
  district_id = as.numeric(as.factor(d_q2$district)),
  age.centered = standardize(d_q2$age.centered),
  living.children = standardize(d_q2$living.children),
  urban=d_q2$urban
)

dat_list_q2 <- list(
  C=dat_q2$use.contraception,
  D=dat_q2$district_id,
  U=dat_q2$urban,
  Ch=dat_q2$living.children,
  A=dat_q2$age.centered,
  N=1934,
  K=max(dat_q2$district_id)
)

```

```{r}
model_q2_1_fit <- stan(file='week09/09_q2_1.stan', data=dat_list_q2, cores=4)
loo(model_q2_1_fit)
```
```{r}
model_q2_2_fit <- stan(file='week09/09_q2_2.stan', data=dat_list_q2, cores=4)
loo(model_q2_2_fit)
```

## Question 3

Modify any models from Problem 2 that contained that children variable and
model the variable now as a monotonic ordered category, like education from the
week we did ordered categories. Education in that example had 8 categories. Children here will have fewer (no one in the sample had 8 children). So modify the code
appropriately. What do you conclude about the causal influence of each additional
child on use of contraception?


```{r}
d_q3 <- bangladesh

d_q3$district_id <- as.integer(as.factor(d_q3$district))

dat_q3 <- data.frame(
  use.contraception=d_q3$use.contraception,
  district_id = as.numeric(as.factor(d_q3$district)),
  age.centered = standardize(d_q3$age.centered),
  living.children = d_q3$living.children,
  urban=d_q3$urban
)

dat_list_q3 <- list(
  C=dat_q3$use.contraception,
  D=dat_q3$district_id,
  U=dat_q3$urban,
  Ch=dat_q3$living.children,
  A=dat_q3$age.centered,
  N=1934,
  K=max(dat_q3$district_id)
)

```


```{r}
model_q3_1_fit <- stan(file='week09/09_q3_1.stan', data=dat_list_q3, cores=4)
loo(model_q3_1_fit)
```

