---
title: "Homework 8"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
library(rethinking)
library(ggplot2)
library(dplyr)
library(tidyr)
library(dagitty)
library(ggdag)
library(rstan)

```


## Question 1
Revisit the Reed frog survival data, data(reedfrogs), and add the predation
and size treatment variables to the varying intercepts model. Consider models with
either predictor alone, both predictors, as well as a model including their interaction.
What do you infer about the causal influence of these predictor variables? Also focus
on the inferred variation across tanks (the  across tanks). Explain why it changes
as it does across models with different predictors included.


```{r}
data(reedfrogs)

d_q1 <- reedfrogs

head(d_q1)


dat_list_q1 <- list(
  S=d_q1$surv,
  D=d_q1$density,
  P=as.integer(d_q1$pred) - 1L,
  Z=as.integer(d_q1$size),
  N=48
)
```



First model surv with only pred alone

```{r}
model_q1_1_fit <- stan(file='week08/08_q1_1.stan', data=dat_list_q1, cores=4)

```
```{r}
model_q1_2_fit <- stan(file='week08/08_q1_2.stan', data=dat_list_q1, cores=4)

```
```{r}
model_q1_3_fit <- stan(file='week08/08_q1_3.stan', data=dat_list_q1, cores=4)
```
```{r}
model_q1_4_fit <- stan(file='week08/08_q1_4.stan', data=dat_list_q1, cores=4)
```
```{r}
model_q1_5_fit <- stan(file='week08/08_q1_5.stan', data=dat_list_q1, cores=4)
```

```{r}
results <- rbind(
  data.frame(model='m5', sigma=rstan::extract(model_q1_5_fit)$A_sigma),
  data.frame(model='m4', sigma=rstan::extract(model_q1_4_fit)$A_sigma),
  data.frame(model='m3', sigma=rstan::extract(model_q1_3_fit)$A_sigma),
  data.frame(model='m2', sigma=rstan::extract(model_q1_2_fit)$A_sigma),
  data.frame(model='m1', sigma=rstan::extract(model_q1_1_fit)$A_sigma)
)

```

```{r}



results %>% 
  ggplot(aes(x=model, y=sigma)) +
  geom_boxplot()+
  coord_flip()+
  scale_y_log10()



```


## Question 2

Now, focus on predicting use.contraception, clustered by district_id. Fit
both (1) a traditional fixed-effects model that uses an index variable for district and
(2) a multilevel model with varying intercepts for district. Plot the predicted proportions of women in each district using contraception, for both the fixed-effects
model and the varying-effects model. That is, make a plot in which district ID is on
the horizontal axis and expected proportion using contraception is on the vertical.
Make one plot for each model, or layer them on the same plot, as you prefer. How
do the models disagree? Can you explain the pattern of disagreement? In particular, can you explain the most extreme cases of disagreement, both why they happen
where they do and why the models reach different inferences?

```{r}

data(bangladesh)

d_q2 <- bangladesh

d_q2$district_id <- as.integer(as.factor(d_q2$district))

dat_list_q2 <- list(
  C=d_q2$use.contraception,
  D=d_q2$district_id,
  N=1934,
  K=max(d_q2$district_id)
)

```
```{r}
model_q2_1_fit <- stan(file='week08/08_q2_1.stan', data=dat_list_q2, cores=4)

```

```{r}
model_q2_2_fit <- stan(file='week08/08_q2_2.stan', data=dat_list_q2, cores=4)
```

```{r}
compare(model_q2_1_fit, model_q2_2_fit)

```
```{r}

post1 <- extract.samples( model_q2_1_fit )
post2 <- extract.samples( model_q2_2_fit )

p1 <- apply( post1$dD, 2 , mean )
p2 <- apply( post2$dDk, 2 , mean )

nd <- max(dat_list_q2$D)
plot( NULL , xlim=c(1,nd) , ylim=c(-1,1) , ylab="prob use contraception" ,
xlab="district" )
points( 1:nd , inv_logit(p1) , pch=16 , col='blue' )
points( 1:nd , inv_logit(p2), col='red')
#abline( h=mean(inv_logit(post2$a_bar)) , lty=2 )

```
## Question 3

Return to the Trolley data, data(Trolley), from Chapter 12. Define and fit a
varying intercepts model for these data. By this I mean to add an intercept parameter
for the individual to the linear model. Cluster the varying intercepts on individual
participants, as indicated by the unique values in the id variable. Include action,
intention, and contact as before. Compare the varying intercepts model and a
model that ignores individuals, using both WAIC/LOO and posterior predictions.
What is the impact of individual variation in these data?

```{r}
data(Trolley)

d_q3 <- Trolley

edu_levels <- c(6, 1, 8, 4, 7, 2, 5, 3)
d_q3$edu_new <- edu_levels[d_q3$edu]

dat_list_q3 <- list(
  response=d_q3$response,
  P=as.numeric(factor(d_q3$id)),
  A=d_q3$action,
  I=d_q3$intention,
  C=d_q3$contact,
  IA=d_q3$intention * d_q3$action,
  IC=d_q3$intention * d_q3$contact,
  E=as.integer(d_q3$edu_new),
  F=1-d_q3$male,
  AGE=standardize(d_q3$age),
  K=7,
  NP=331,
  N=9930
) 
```

```{r}
model_q3_1_fit <- stan(file='week08/08_q3_1.stan', data=dat_list_q3, cores=4)
```
```{r}
model_q3_2_fit <- stan(file='week08/08_q3_2.stan', data=dat_list_q3, cores=4)
```

```{r}
loo(model_q3_1_fit)
```

```{r}
loo(model_q3_2_fit)
```

```{r}
compare(model_q3_1_fit, model_q3_2_fit)
```


