---
title: "Homework 1"
output:
  pdf_document: default
---

Used libraries are:

 \begin{itemize}
  \item rethinking
  \item ggplot2
\end{itemize}

```{r setup, include=FALSE}
library(rethinking)
library(ggplot2)
```



## Question 1

Suppose the globe tossing data had turned out to be 8 water in 15 tosses.
Construct the posterior distribution, using grid approximation. Use the
same flat prior as before.


```{r}
p_grid <- seq(0, 0.99, 1/100)
prior_of_p <- rep(1/100, 100)

likelihood <- dbinom(8, 15, p_grid) * prior_of_p
posterior <- likelihood / sum(likelihood)

df <- data.frame(p=p_grid, posterior=posterior)

ggplot(df, aes(x=p, y=posterior)) + geom_line() 


```

```{r}
samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE) 

pi <- PI(samples, 0.98)
posterior_mean <- mean(samples)

```


Posterior mean is around `r posterior_mean` and 99 interval is between `r pi['1%']` and `r pi['99%']`


## Question 2

Start over in 1, but now use a prior that is zero below p = 0.5 and a constant above p = 0.5.
This corresponds to prior information that a majority
of the Earth’s surface is water. What difference does the better prior make?
If it helps, compare posterior distributions (using both priors) to the true
value p = 0.7.

```{r}

p_grid <- seq(0, 1, 1/100)
prior_of_p_blow_05 <- rep(0, 51)
prior_of_p_above_05 <- rep(1/50, 50)

prior_of_p <- c(prior_of_p_blow_05, prior_of_p_above_05)
names(prior_of_p) <- p_grid


likelihood <- dbinom(8, 16, p_grid) * prior_of_p
posterior <- likelihood / sum(likelihood)

df <- data.frame(p=p_grid, posterior=posterior)

ggplot(df, aes(x=p, y=posterior)) + geom_line() 


```

```{r}
samples_2 <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE) 

pi_2 <- PI(samples_2, 0.98)
posterior_mean_2 <- mean(samples_2)

```


Posterior mean is around `r posterior_mean_2` and 99 interval is between `r pi_2['1%']` and `r pi_2['99%']`


```{r}

dens( samples , xlab="p" , xlim=c(0,1) , ylim=c(0,8) , show.HPDI=0.5)
dens( samples_2 , add=TRUE , lty=2 , show.HPDI=0.5)


```



## Question 3


This problem is more open-ended than the others. Feel free to collaborate on the solution. Suppose you want to estimate the Earth’s proportion of
water very precisely. Specifically, you want the 99% percentile interval of the
posterior distribution of p to be only 0.05 wide. This means the distance between the upper and lower bound of the interval should be 0.05. How many
times will you have to toss the globe to do this? I won’t require a precise
answer. I’m honestly more interested in your approach.


```{r}

f <- function(trial){
  
  p_grid <- seq(0, 1, 1/100)
  prior_of_p <- rep(1/101, 101)
  
  names(prior_of_p) <- p_grid

  
  likelihood <- dbinom(trial/2, trial, p_grid) * prior_of_p
  posterior <- likelihood / sum(likelihood)
  
  samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE) 
  
  interval <- PI(samples, 0.98)
  
  return(as.numeric(interval['99%'] - interval['1%']))
}

Nlist <- c( 20 , 50 , 100 , 200 , 500 , 1000 , 2000 )
Nlist <- rep( Nlist , each=100 )

width <- sapply( Nlist , f )

plot( Nlist , width )
abline( h=0.05 , col="red" )
#ggplot(df, aes(x=trial, y=percentile_interval)) + geom_line() 


```

```{r}

df <- data.frame(width=tapply(width, Nlist, FUN=mean))

knitr::kable(df, floating.environment="sidewaystable")

```

