---
title: "Homework 2"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

Used libraries are:

```{=tex}
\begin{itemize}
  \item rethinking
  \item ggplot2
\end{itemize}
```
```{r setup, include=FALSE}
library(rethinking)
library(ggplot2)
```

## Question 1

**The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% compatibility intervals for each of these individuals. That is, fill in the table below, using model-based predictions**

```{r}

data(Howell1)

d <- Howell1
d2 <- d[d$age >= 18, ]

```

Lets first visualize the prior predictive simulation

```{r}

N <- 100
a <- rnorm(N, 178, 20)
b <- rlnorm(N, 0, 1)

plot(height~weight, data=d2, xlim=range(d2$weight), ylim=c(-100, 400), xlab="weight", ylab="height")

abline(h=0, lty=2)
abline(h=272, lty=1, lwd=0.5)

xbar <- mean(d2$weight)

for(i in 1:N) curve(a[i] + b[i] * (x-xbar), from=min(d2$weight), to=max(d2$weight), add=TRUE, col=col.alpha("black", 0.2))
```

Prior predictive simulation makes sense, lets get posterior distribution

```{r}


flist <- alist(
  height ~ dnorm(mu, sigma),
  mu ~ a + b * (weight - xbar),
  a ~ dnorm(178, 20),
  b ~ dlnorm(0, 1),
  sigma ~ dunif(0, 50)
)


homework_1_model <- quap(flist, data=d2)

```

Now we have the model, lets simulate posterior predictive distributions


```{r}
new_data <- data.frame( weight=c(45,40,65,31,53) )

simulated <- sim(homework_1_model, data=new_data, n=1e4)


```


```{r}

df <- data.frame(apply(simulated, 2, PI))

knitr::kable(df, floating.environment="sidewaystable")

```

## Question 2 and 3

**Model the relationship between height (cm) and the natural logarithm of
weight (log-kg): log(weight). Use the entire Howell1 data frame, all 544
rows, adults and non-adults. Use any model type from Chapter 4 that you
think useful: an ordinary linear regression, a polynomial or a spline. Plot
the posterior predictions against the raw data.**

**Plot the prior predictive distribution for the polynomial regression model
in Chapter 4. You can modify the the code that plots the linear regression
prior predictive distribution. 20 or 30 parabolas from the prior should suffice to show where the prior probability resides. Can you modify the prior
distributions of α, β1, and β2 so that the prior predictions stay within the
biologically reasonable outcome space? That is to say: Do not try to fit the
data by hand. But do try to keep the curves consistent with what you know
about height and weight, before seeing these exact data**


Lets assume the relationship between weight and height is weight = height^3

```{r}
d3 <- as.data.frame(scale(d))
d3$height <- d3$height - min(d3$height)
```


```{r}

N <- 100
a <- rnorm(N, -2, 0.1)
b <- rnorm(N, 0.1, 0.1)
c <- rnorm(N, 0.05, 0.05)
e <- rnorm(N, 0.05, 0.05)

plot(weight~height, data=d3, ylim=range(d3$weight), xlim=range(d3$height), ylab="weight", xlab="height")

x_min <- min(d3$height)

for(i in 1:N) curve(a[i] + + b[i] * x + c[i] * x^2 + e[i] * x^3, from=min(d3$height), to=max(d3$height), add=TRUE, col=col.alpha("black", 0.2))

```

```{r}
flist <- alist(
  weight ~ dnorm(mu, sigma),
  mu ~ a + b *  height + c * height ^ 2 + e * height ^ 3,
  a ~ dnorm(-2, 1),
  b ~ dnorm(0.1, 0.1),
  c ~ dnorm(0.05, 0.05),
  e ~ dnorm(0.05, 0.05),
  sigma ~ dunif(0, 100)
)

homework_model_2 <- quap(flist, data=d3)

```

```{r}

prior <- extract.prior( homework_model_2 )

w_seq <- seq( from=min(d3$height) , to=max(d3$height), length.out=50 )

mu <- link( homework_model_2 , post=prior, data=list( height= w_seq))

plot( NULL , xlim=range(w_seq) , ylim=range(d3$weight), xlab="weight (std)" , ylab="height" )
for ( i in 1:50 ) lines( w_seq , mu[i,] , col=col.alpha("black",0.5) )

```


```{r}
height_seq <- seq(min(d3$height), max(d3$height), length.out=30)
simulated <- sim(homework_model_2, data=data.frame(height=height_seq), n=1e4)

height.PI <- apply(simulated, 2, mean)
plot(weight~height, data=d3,  xlab="weight", ylab="height")
lines(height_seq, height.PI)

```
```{r}
d$log_weight <- log(d$weight)

m2 <- quap(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b* log_weight ,
a ~ dnorm( 178 , 20 ) ,
b ~ dlnorm( 0 , 1 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data=d )
plot( d$weight , d$height , col=col.alpha(rangi2,0.7) )
x_seq <- log(1:60)
mu <- sim( m2 , data=list(log_weight=x_seq) )
mu_mean <- apply(mu,2,mean)
mu_ci <- apply(mu,2,PI,0.99)
lines( exp(x_seq) , mu_mean )
shade( mu_ci , exp(x_seq) )

```

```{r}

posterior <- extract.samples( m2 )

w_seq <- seq( from=min(d$weight) , to=max(d$weight), length.out=50 )

mu <- link( m2, post=prior, data=list( weight= w_seq, log_weight=log(w_seq)))

plot( NULL , xlim=range(w_seq) , ylim=range(d$height), xlab="weight (std)" , ylab="height" )
for ( i in 1:50 ) lines( w_seq , mu[i,] , col=col.alpha("black",0.5) )
```





